---
title: "STATS5099 Data Mining and Machine Learning"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

```{r setup, include = FALSE}
library(webexercises)
library(tinytex)
library(dplyr)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
library(reticulate)
```


```{r include=FALSE, echo=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


# Welcome to DMML Lab 9

In week 9, we have studied deep learning models to handle sequential data, specifically recurrent neural network (RNN), long short-term memory (LSTM), gated recurrent unit (GRU) and transformer. In this lab, we will see how to fit a time series using RNN and illustrate the sentiment analysis example in lecture note in more details. All codes are implemented in Python. 

```{python, echo=FALSE}
import warnings
warnings.filterwarnings("ignore", category=DeprecationWarning)
```

